---
title: "Assignment 4 Planning and Modelling of Public Transport"
author: "Vilém Knap"
date: "April 22, 2021"
output: 
  pdf_document:
    fig_caption: yes
header-includes:
 \usepackage{float}
 \floatplacement{figure}{H}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Loading the dataset

```{r}
 dataset <- read.csv("~/DTU/3. semestr/Planning and Modelling of Public Transport/Assignment 4/dataset.csv", sep=";")
```

# Descriptive statistics

## Overview of the data

```{r}
## Dimension of HE (number of rows and columns)
dim(dataset)
```


```{r}
## Column names
names(dataset)
```

```{r}
## The first rows
head(dataset)
```

```{r}
## The last rows
tail(dataset)
```

```{r}
## Default summary
summary(dataset)
```

```{r}
## Another summary function also including the data type
str(dataset)
```


```{r}
sapply(dataset, function(x) sum(is.na(x)))
```

## Handling with NA values
### Income
As the dataset contians quite low amount of samples (215) we have decided to impute missing values. Regarding the income we have missing observations for students only, so we will use most frequent value for students which is 7000DKK for lower and 10000DKK for upper range


```{r}
dataset$Income_monthly_lower[is.na(dataset$Income_monthly_lower)] <- 7000
```

```{r}
dataset$Income_monthly_upper[is.na(dataset$Income_monthly_upper)] <- 10000
```

### Work or study distance
Regarding the distances we have 13 observations for upper and 6 for lower range.
It seems to be appropriate approach to impute distances based on region of living and occuptaion

```{r}
new.data <- dataset[ is.na(dataset$work_or_study_distance_lower | dataset$work_or_study_distance_upper)  & (dataset$Occupation == 5)  , ]
```
As the unemployed people do not commute to any workplace, we fill their work or study distance with zero

```{r}
dataset$work_or_study_distance_lower[is.na(dataset$work_or_study_distance_lower)] <- 0
dataset$work_or_study_distance_upper[is.na(dataset$work_or_study_distance_upper)] <- 0
```

## Deleting X column

```{r}
dataset_filtered <- subset(dataset, select = -X )
```


## Histograms

```{r}
hist(dataset_filtered$Vaccinated_now, breaks = 2)
hist(dataset_filtered$Vaccinated_future, breaks = 2)
hist(dataset_filtered$how_often_PT_before_covid, breaks = 6)
hist(dataset_filtered$how_often_PT_during_lockdown_covid, breaks = 6)
hist(dataset_filtered$how_often_PT_no_lockdown_covid, breaks = 6)
#Safety on the bus
hist(dataset_filtered$safety_bus_no_vacc, main="How safe did you feel on the bus?", breaks = 6)
hist(dataset_filtered$safety_bus_now_11percent_vacc, main="How safe are you feeling on the bus?", breaks = 6)
hist(dataset_filtered$safety_bus_40percent_vacc, main="How safe would you feel in the bus when 40% population vaccinated?", breaks = 6)
hist(dataset_filtered$safety_bus_75percent_vacc, main="How safe would you feel in the bus when 75% population vaccinated?", breaks = 6)
hist(dataset_filtered$safety_bus_95percent_vacc, main="How safe would you feel in the bus when 95% population vaccinated?", breaks = 6)
#Safety on the train
hist(dataset_filtered$safety_train_no_vacc, main="How safe did you feel on the train?", breaks = 6)
hist(dataset_filtered$safety_train_now_11percent_vacc, main="How safe are you feeling on the train?", breaks = 6)
hist(dataset_filtered$safety_train_40percent_vacc, main="How safe would you feel in the train when 40% population vaccinated?", breaks = 6)
hist(dataset_filtered$safety_train_75percent_vacc, main="How safe would you feel in the train when 75% population vaccinated?", breaks = 6)
hist(dataset_filtered$safety_train_95percent_vacc, main="How safe would you feel in the train when 95% population vaccinated?", breaks = 6)
#Safety on the metro
hist(dataset_filtered$safety_metro_no_vacc, main="How safe did you feel on the metro?", breaks = 6)
hist(dataset_filtered$safety_metro_now_11percent_vacc, main="How safe are you feeling on the metro?", breaks = 6)
hist(dataset_filtered$safety_metro_40percent_vacc, main="How safe would you feel in the metro when 40% population vaccinated?", breaks = 6)
hist(dataset_filtered$safety_metro_75percent_vacc, main="How safe would you feel in the metro when 75% population vaccinated?", breaks = 6)
hist(dataset_filtered$safety_metro_95percent_vacc, main="How safe would you feel in the metro when 95% population vaccinated?", breaks = 6)
```

## More histograms

```{r}
#Factors importance during reopening
hist(dataset_filtered$more_frequent_vehicle_desinfection_importance_reopening, main="How important do you feel is more frequent desinfection of vehicles during the reopening?", breaks = 6)

hist(dataset_filtered$increased_cleanliness_importance_reopening, main="How important do you feel is increased cleanliness during the reopening?", breaks = 6)

hist(dataset_filtered$decreased_ticket_price_importance_reopening, main="How important do you feel is to decrease ticket price during the reopening?", breaks = 6)

hist(dataset_filtered$more_frequent_service_importance_reopening, main="How important do you feel is more frequent service during the reopening?", breaks = 6)

hist(dataset_filtered$reduced_capacity_importance_reopening, main="How important do you feel is decreased capacity of vehicles during the reopening?", breaks = 6)

hist(dataset_filtered$reduced_time_spent_in_PT_importance_reopening, main="How important do you feel is to reduced time spent in vehicles during the reopening?", breaks = 6)

hist(dataset_filtered$mask_usage_in_PT_importance_reopening, main="How important do you feel is to use mask in vehicles during the reopening?", breaks = 6)

hist(dataset_filtered$opened_windows_in_vehicle_importance_reopening, main="How important do you feel is to have windows opened in vehicles during the reopening?", breaks = 6)

hist(dataset_filtered$X2_meters_distance_in_PT_importance_reopening, main="How important do you feel is to maintain 2 meters distance in vehicles during the reopening?", breaks = 6)

hist(dataset_filtered$online_PT_schedule_importance_reopening, main="How important do you feel is to have online schedule during the reopening?", breaks = 6)

hist(dataset_filtered$increased_fare_percieving, main="How would you percieve increasing of fare during the reopening?", breaks = 6)

hist(dataset_filtered$reduced_seat_percieving, main="How would you percieve reduced seats in public transport vehicles during reopening?", breaks = 6)
## IS THIS SAME QUESTION AS REDUCED CAPACITY?

hist(dataset_filtered$reduced_private_vehicles_percieving, main="How would you percieve to reduce amount of private vehicles during reopening?", breaks = 6)
## How would you achieve that, more taxes on cars? 

hist(dataset_filtered$bus_fleet_electrification_increased_fare_percieving, main="How would you percieve to electrify bus fleet with corresponding increase of fare?", breaks = 6)
```

### Travel times and cost
```{r}
hist(dataset_filtered$car_travel_cost, main="Car travel cost empirical distribution", breaks = 50)
hist(dataset_filtered$car_travel_time, main="Car travel time empirical distribution", breaks = 10)

hist(dataset_filtered$car_share_walk_cost, main="Car share travel cost empirical distribution", breaks = 10)
hist(dataset_filtered$car_share__walk_time, main="Car share travel time empirical distribution", breaks = 10)

hist(dataset_filtered$bus_travel_cost, main="Bus travel cost empirical distribution", breaks = 10)
hist(dataset_filtered$bus_travel_time, main="Bus travel time empirical distribution", breaks = 10)

hist(dataset_filtered$bicycle_travel_cost, main="Bicycle travel cost empirical distribution", breaks = 10)
hist(dataset_filtered$bicycle_travel_time, main="Bicycle travel time empirical distribution", breaks = 10)

hist(dataset_filtered$cycling_train_travel_cost, main="cycling_train travel cost empirical distribution", breaks = 10)
hist(dataset_filtered$cycling_train_travel_time, main="cycling_train travel time empirical distribution", breaks = 10)

```
Uniformly distributed...

### Acess to car

```{r}
hist(dataset_filtered$Access_to_car,main="Acess to car",breaks = 2)
```

We have only respondents without access to car. That can be caused by selection bias as the questionnare was spread out on social networks and most of the respondents are studenets.
### Stated mode choice
```{r}
hist(dataset_filtered$stated_mode_choice_11._vaccinated,main="Stated mode choice when 11 percent of population vaccinated", breaks = 6)
hist(dataset_filtered$stated_mode_choice_40._vaccinated,main="Stated mode choice when 40 percent of population vaccinated", breaks = 6)
hist(dataset_filtered$stated_mode_choice_75._vaccinated,main="Stated mode choice when 75 percent of population vaccinated", breaks = 6)
hist(dataset_filtered$stated_mode_choice_95._vaccinated,main="Stated mode choice when 95 percent of population vaccinated", breaks = 6)
```
We can see decreasing trend of car usage when vaccination of population increases. On the beginning 80 respondents has chosen car as their mean of travel from DTU to Rosenborg Slot. If 40 percent of population is vaccinated, approx 65 respondents would go by car. If 75 percent of population is vaccinated approx 50 respondents would go by car. If 95 percent of population is vaccinated approx 49 respondents would go by car.
The car sharing options is the least favorite when 11, 75 and 95 percent of population is vaccinated. 
With increasing amount of vaccinated population combination of bike and train increases. If 75 percent and or more people vaccinated, then it becomes the most favorite option. 
In the beginning (11 percent vaccinated population) bus is the least favorite option. As the vaccination proceeds hypothetical usage of bus increases. If 95% population is vaccinated, it become the second most favorite option, after combination of bike and train.

### Socio-demographic characteristics
```{r}
hist(dataset_filtered$Gender, main="Distribution of gender", breaks=3)
hist(dataset_filtered$Age_group, main="Distribution of age", breaks=6)
hist(dataset_filtered$Household_size, main="Distribution of household size", breaks=6)
hist(dataset_filtered$Occupation, main="Distribution of occupations", breaks=6)
```
### Geographic characteristics

```{r}
hist(dataset_filtered$work_or_study_distance_lower, breaks = 6, main="Distribution of distances")
hist(dataset_filtered$work_or_study_distance_upper, breaks = 6, main="Distribution of distances")
```

### Economic characterstics

```{r}
hist(dataset_filtered$Income_monthly_lower, breaks = 6, main="Distribution of income lower range")
hist(dataset_filtered$Income_monthly_upper, breaks = 6, main="Distribution of income upper range")
```

We can observe that we have slightly more observations for female than male.
The majority of the respondest were young people.

###

## Box plots

## Pairs plot

```{r}
par(mfrow=c(2,2))
plot(lm_cubic_root)
```


## Covariance matrix 
First we need to exlude region of living to have numerical variables only.

```{r}
dataset_filtered_numerical <- subset(dataset_filtered, select=-c(Region_of_living))
```

```{r}
cov(dataset_filtered_numerical)
```

## Correlation matrix 
```{r}
cor(dataset_filtered_numerical)
```







